---
title: "Mdhumne_hm03"
author: "Mrinmayee Dhumne"
date: "4/1/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
 ****Use this dataset to estimate the volatility of the efficient security price.
```{r}

library(xts)
library(highfrequency)
Sys.setenv(TZ = "GMT")  # work in East Coast Time Zone
options(digits.secs=3)
load("sampleTQdata.RData")

tqdata<- tqdata
head(tqdata)
tail(tqdata)
length(tqdata$SIZE)    #8153 trades 



# Plot prices 

asks <- as.numeric(tqdata$OFR)
bids <- as.numeric(tqdata$BID)
mids <- 0.5*bids + 0.5*asks

qspread <- asks - bids

pmin = min(as.numeric(tqdata$PRICE))
pmax = max(as.numeric(tqdata$PRICE))
plot(as.numeric(tqdata$PRICE),col="red", type="l", ylab="Trade price", 
     xlab="Trade #", main="Trade price ", ylim=c(pmin-0.1,pmax+0.1))
lines(mids, type="l", col="blue")

plot(as.numeric(tqdata$SIZE),col="pink", type="l", 
     ylab="Trade size", 
     xlab="Trade #", main="Trade volume", ylim=c(0,10000))
```

Problem 1. Calibrate the Roll model on the time series of trade prices pt, and estimate the Roll model parameters c,σu. What is the estimated bid-ask spread 2c?

```{r}
#calibrate the Roll model on the real data
pt <- as.numeric(tqdata$PRICE)
dpt <- diff(pt)  # price change

head(dpt)

covpt <- acf(dpt, lag.max=20, type="covariance", 
             main="covariance of price changes")
# Roll model estimate of bid-ask spread
gamma0 <- covpt$acf[1]
print(gamma0)       

gamma1 <- covpt$acf[2]
print(gamma1)     

```
Gamma 0 = 0.002884046
Gamma 1 = -0.0004072397
```{r}
#the Roll model parameters c,σu
c_param <- sqrt(-covpt$acf[2])
print(c_param)    # c = 0.02018018

# Calculating Sigma u
Sig2u <- gamma0 +2*gamma1   #σu2
Sigu <- sqrt(Sig2u)             # σu
print(Sigu)                # σu = 0.04549248 

n.trades <- length(pt)
sigann <- sqrt(252*n.trades)*Sigu
print(sigann)         #65.20767
sav <-mean(pt)
sigannln <- sigann/sav
print(sigannln)        #0.3412752
```
c  = 0.02018018
σu = 0.04549248 

Problem2.Repeat the calibration, using this time the time series of the log-trade prices log pt. What are the parameters?
```{r}
pt <- as.numeric(tqdata$PRICE)
log_pt <- log(pt)
log_dpt <- diff(log_pt)
log_acpt <- acf(log_dpt, lag.max =20, type= "correlation", plot =TRUE, main="Autocorrelation")

covpt <- acf(log_dpt, lag.max=20, type="covariance", 
             main="covariance of price changes")


#Roll model estimate of bid-ask spread

log_gamma0 <- sd(dpt)^2
print(log_gamma0)

log_gamma1 <- covpt$acf[2]
print(gamma1)     #-1.116191e-08

#Parameter C of the roll model
c_param <- sqrt(-covpt$acf[2])
print(c_param)    # c = 0.0001056499

# Calculating Sigma u
Sig2u <- gamma0 +2*gamma1   #sigma square
Sigu <- sqrt(Sig2u)             # sigma u
print(Sigu)                # sigma u = 0.0002380327 

n.trades <- length(pt)
sigann <- sqrt(252*n.trades)*Sigu
print(sigann)         #0.341189 
sav <-mean(pt)
sigannln <- sigann/sav
print(sigannln)        #0.00178567
```
Gamma 0 = 7.89737e-08
Gamma 1 = -1.116191e-08
c  = 0.0001056499
σu = 0.0002380327

Problem 3.1. Compute the signature plot showing σ2 (q) for lags q = 1 : 100. Recall day that σday(q) is the daily price volatility measured by keeping only every q-th trade price. For this part use the function realizedVar(q) as define.
```{r}
realizedVar <- function(q){
  pr<- as.numeric(tqdata$PRICE)
  rCov(diff(pr, lag=q, differences=1)/q)}
# compute the signature plot sigma.day(q) = sqrt(RV(q))
sig_data <- NULL
for(q in 1:100){
  sig_data <- c(sig_data, sqrt(realizedVar(q)))
}
plot(sig_data, type ="l", main="Signature plot")
abline(h= sqrt(realizedVar(50)), col= "purple")
```
Problem 3.2.Comment on the shape of the signature plot. What do we learn from it?

The volatility signature plot gives us a useful insigt abut bias of realised variance estimate is manifesting itself in an upward sloping pattern
as the samppling interval or lag becomes shoter that is bias increases with assumed availability of high frequency price
here sampling interval is time period divided by assumed available high frequency

Problem 4.Compare the daily volatility obtained from point 1. σ2 = ntradesσu2
day,Roll
with the daily volatility obtained in point 3 with a lag q5min corresponding
to a 5 minute interval between trades, assuming that the rate of trading is constant during the day.
First determine the lag q5min = ntrades 5 and then compute the volatility.
```{r}

n.trades <- length(tqdata$PRICE)

sig_p2_daily <- n.trades * Sig2u
sig_p_daily <- sqrt(sig_p2_daily)
sig_p_daily       #0.02149108


q5mins <- (n.trades*5)/390
rv5 = realizedVar(q5mins)
rv5        ##0.1850623

sig5min_daily =sqrt(rv5)
sig5min_daily     #0.43018
```
The daily volatality from roll model = 0.02149108
Daily volatality from lag q5min= 0.1850623
The daily volatality calculated using q5mins is higher than the daily volatality calculated using roll model. Since, Roll model assumes uncorrelated trade indicatores and its more accurate.



